---
title: "Data607 Sentiment Analysis"
author: "Nicholas Kunze"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if("tidytext" %in% rownames(installed.packages()) == FALSE) {install.packages("tidytext")}
if("textdata" %in% rownames(installed.packages()) == FALSE) {install.packages("textdata")}
```

## Overview

The base code for this assignment is originally from "Text Mining with R: A Tidy Approach" by Julia Silge and David Robinson, Chapter 2: <https://www.tidytextmining.com/sentiment.html#sentiment>

This assignment focuses on sentiment analysis. To quote the original text, "We can use the tools of text mining to approach the emotional content of text programmatically".

## Base Code

```{r afinn-1}
library(tidytext)

get_sentiments("afinn")
```

```{r afinn-2}
get_sentiments("bing")
```

```{r afinn-3}
get_sentiments("nrc")
```

```{r base-4}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r base-5}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r base-6}
library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

```{r base-7}
library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## New Corpus

As a lover of Tolkien, I'm curious on the sentiment analysis behind his most popular, "The Lord of the Rings". Lets add the loughran sentiment lexicon and use that for our analysis. I've pulled the text from an existing source I found on github.

```{r new-corpus}
library(tibble)
library(RCurl)
lotr <- as.tibble(getURI("https://raw.githubusercontent.com/wess/iotr/master/lotr.txt"))
lotrWords <- lotr %>% unnest_tokens(word,value) %>% anti_join(stop_words)
head(lotrWords)
```

```{r lotr-check}
loughran_pos <- get_sentiments("loughran") %>% 
  filter(sentiment == "positive")
lotrWords %>% 
  inner_join(loughran_pos) %>%
  count(word, sort = TRUE)
```

```{r lotr-posvsneg}
lotrWords$linenumber <- 1:nrow(lotrWords)
lotrSentiment <- lotrWords %>% 
  inner_join(get_sentiments("loughran")) %>%
  count(index = linenumber %/% 500, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
ggplot(lotrSentiment, aes(index, sentiment)) +
  geom_col(show.legend = FALSE)
```

I honestly would have expected more positive sentiment towards the beginning, though I suppose less negative sentiment will have to do. Compared to Jane Austen, Tolkien is much more negative...
